# Analysing the analysers - the sequel

```{r preliminaries}
# run following two lines from the CLI; will then generate the rest of the HTML file
# setwd("dssurvey")
# library("knitr"); knit2html("analyse.Rmd")
source("readdata.R"); mydata <- readdata("data/Gegevens_Allemaal_150413.zip")
source("geolocate.R")
library(ggplot2, quietly=TRUE)
library(ggmap, quietly=TRUE)
library(dplyr, quietly=TRUE)
library(reshape2, quietly=TRUE)
```

## Number of responses
How many responses did we get, and how complete were they?

```{r numberresponses}
dat <- tbl_df(mydata)
completed <- matrix(rep(0, 15*nrow(mydata)), ncol=15)
completed[, 1] <- !is.na(dat$Age)
completed[, 2] <- !is.na(dat$Gender)
completed[, 3] <- dat$OrganisationType!="" | dat$OrganisationTypeOther==""
completed[, 4] <- !is.na(dat$OrganisationSize)
completed[, 5] <- dat$HierarchyOrganisation!=""
completed[, 6] <- apply((select(dat, starts_with("Role"))!=""), 1, sum)>0
completed[, 7] <- is.na(dat$OrganisationIndustry)!="" & dat$OrganisationIndustryOther==""
completed[, 8] <- dat$Degree!="" | dat$DegreeOther!=""
completed[, 9] <- apply((select(dat, starts_with("Studies"))!=""), 1, sum)>0
completed[, 10] <- !is.na(dat$YearsExperience)
completed[, 11] <- apply((select(dat, starts_with("Skills"))!=""), 1, sum)>0
completed[, 12] <- apply((select(dat, starts_with("Type"))!=""), 1, sum)>0
completed[, 13] <- apply((select(dat, starts_with("Motivation"))!=""), 1, sum)>0
completed[, 14] <- apply((select(dat, starts_with("Activities"))!=""), 1, sum)>0
completed[, 15] <- apply((select(dat, starts_with("Training"))!=""), 1, sum)>0
completed[is.na(completed[, ])] <- 0
colnames(completed) <- c("Age", "Gender", "OrganisationType", "OrganisationSize", 
		"Hierarchy", "Role", "Industry", "Degree", "Studies", "Experience",
		"Skills", "Type", "Motivation", "Activities", "Training")
# R classic plot
# barplot(apply(completed, 2, sum), las=2)

# ggplot2
tmp <- melt(completed)
tmp <- tmp[tmp$value!=0, ]
names(tmp)[2] <- "question" 
ggplot(tmp, aes(question)) + 
		geom_bar() +
		theme(axis.text.x=element_text(angle=90, hjust=1))
```

There is a downward trend, but not too bad, in spite of the length of the survey.

## Time and date
```{r time}
## R classic plot
#hist(as.numeric(strftime(mydata$StartDate, format="%H")),
#		breaks=0:24, main="Time of day", 
#		xlab="time", xlim=c(0, 24), include.lowest=FALSE, right=FALSE)

# time of day the survey was started - ggplot
timeOfDay <- as.data.frame( as.numeric(strftime(mydata$StartDate, format="%H")))
names(timeOfDay) <- c("Time of Day")
ggplot(data=timeOfDay, aes(x=`Time of Day`)) + geom_histogram(binwidth=1)
```

Note - time is in GMT, so one or two hours different from the time on the clock.
Need to do transform and derive time on the clock if we want to say something about 
behaviour of Data Scientists.

```{r date}
day <- as.numeric(strftime(mydata$StartDate, format="%d"))
month <- as.numeric(strftime(mydata$StartDate, format="%m"))
date <- day + ifelse(month==4, 31, 0)
#hist(date,
#		breaks=10:47, main="Day of month", 
#		xlab="day", include.lowest=FALSE, right=FALSE,
#		xaxt="n")
#axis(1, at=10:47, labels=c(10:31, 1:16))

# cumulative plot
plot(mydata$StartDate[order(mydata$StartDate)], 1:length(mydata$StartDate), 
		xlab="Date & time response received", ylab="Cumulative number of replies", type="p")

# date - ggplot2
dayOfMonth <- as.data.frame(date)
names(dayOfMonth) <- c("Day of Month")
ggplot(data=dayOfMonth, aes(x=`Day of Month`)) + 
		geom_histogram(binwidth=1) +
		scale_x_continuous(breaks=10:47, labels=c(10:31, 1:16))
```

## Where do the responses come from?
Will show a map of geolocation of IP addresses
```{r location, cache=TRUE}
#pos <- geoip(as.character(mydata$IP.Address)) # function from geolocate.R
#pos <- pos[, c("latitude", "longitude")]
#get_belgium <- get_map(c(4.5,51), zoom=7)

#belgium <- ggmap(get_belgium) # actually shows Belgium, Holland and north of France (south to Paris)
#belgium + 
#		geom_point(aes(x=longitude, y=latitude), data=pos, col="red")
```


## Age and gender
```{r gender_age}
counts <- table(mydata$Age)
op <- par(mar = c(12,4,4,2) + 0.1)
barplot(counts, las=2)

counts <- table(mydata$Gender)
op <- par(mar = c(8 ,4,4,2) + 0.1)
barplot(counts)
responded <- sum(counts)
skipped <- nrow(mydata)-responded

op <- par(mar = c(4 ,8,4,2) + 0.1)
# ggplot: problems labeling the y axis
#tmp <- data.frame(Gender=mydata$Gender, Age=as.numeric(mydata$Age))
#ggplot(data=tmp, aes(x=Gender, y=Age))+geom_boxplot()
#	+scale_y_continuous(labels=levels(mydata$Age))
```

Responded: `r responded`; skipped: `r skipped`

## Organisation
```{r organisation}
counts <- table(mydata$OrganisationType)
counts <- counts[-1]
counts <- counts[c(1, 2, 3, 4, 5, 7, 6)]
names(counts)[7] <- "Other"
op <- par(mar = c(12,4,4,2) + 0.1)
barplot(counts, las=2)
par(op)
responded <- sum(counts)
skipped <- nrow(mydata)-responded
```

Responded: `r responded`; skipped: `r skipped`

```{r organisationsize}
d <- as.numeric(as.character(mydata$OrganisationSize))
d <- d[!is.na(d)]; d <- d[d!=1234567]
hist(log(d), xlab="log(Organisation size)", main="Organisation size")
# axis(1, exp(r$breaks), las=2)
responded <- length(d); skipped=nrow(mydata)-responded
```

Responded: `r responded`; skipped: `r skipped`

## Responsibilities and role
```{r responsibilities}
op <- par(mar = c(12,4,4,2) + 0.1)
counts <- table(mydata$HierarchyOrganisation)
names(counts) <- c("", "First-line manager", "Middle-level mamager", "No management resp.", "Top level")
counts <- counts[-1]
counts <- counts[c(1, 2, 4, 3)]
barplot(counts, las=2)
responded <- sum(counts); skipped <- nrow(mydata)-responded
```

Responded: `r responded`; skipped: `r skipped`

```{r role}
roles <- mydata[, names(mydata)[grepl("^Role", names(mydata))]]
roles[is.na(roles$RoleOther), "RoleOther"] <- ""
counts <- rep(0, ncol(roles))
for(i in 1:ncol(roles)) counts[i] <- sum(roles[, i]!="") 
names(counts) <- gsub("^Role", "", names(roles))
names(counts)[5] <- "R & D"
op <- par(mar = c(8,4,4,2) + 0.1)
barplot(counts, las=2)
counts <- rep(0, nrow(roles))
for(i in 1:nrow(roles)) counts[i] <- sum(roles[i, ]!="")
responded <- sum(counts>0); skipped <- nrow(mydata)-responded
```

Responded: `r responded`; skipped: `r skipped`
```{r industry}
counts <- table(mydata$OrganisationIndustry)
counts <- counts[-1]
names(counts)[1] <- "Communications, etc."
names(counts)[6] <- "Healthcare, etc."
names(counts)[10] <- "Other"
counts <- counts[c(1:9, 11:13, 10)]
op <- par(mar = c(12,4,4,2) + 0.1)
barplot(counts, las=2)
responded <- sum(counts); skipped <- nrow(mydata)-responded
```

Responded: `r responded`; skipped: `r skipped`

## Education
```{r degree}
degreeLevels <- c("", "No higher education", "Bachelor", "Master", "Master after Master", "PhD", "Other (please specify)")
tmp <- as.character(mydata$Degree)
mydata$Degree <- factor(tmp, levels=degreeLevels, ordered=TRUE)
levels(mydata$Degree) <- c("", "No higher education", "Bachelor", "Master", "Master after Master", "PhD", "Other")
counts <- table(mydata$Degree)
counts <- counts[-1]
barplot(counts, las=2)
responded <- sum(counts); skipped <- nrow(mydata)-responded
```

Responded: `r responded`; skipped: `r skipped`
```{r studytypes}
studiesLevels <- c("StudiesScience", "StudiesEngineering", "StudiesICT", "StudiesCommerce", "StudiesSocial", "StudiesLaw", "StudiesOther")
studiesCounts <- rep(0, length(studiesLevels))
for(i in 1:length(studiesLevels)) studiesCounts[i] <- sum(mydata[, studiesLevels[i]]!="")
studiesNames <- c("Science or Math", "Engineering", "ICT", "Commerce", "Social", "Law", "Other")
names(studiesCounts) <- studiesNames
barplot(studiesCounts, las=2)
```

## Experience
```{r experience}
hist(mydata$YearsExperience)
```

## Skills
```{r skills}
# names(mydata)[32:53]
skills <- mydata[, names(mydata)[grepl("^Skills", names(mydata)) & !grepl("Other$", names(mydata))]]
names(skills) <- sub("^Skills", "", names(skills))
test1 <- apply(skills, 1, var)
skills <- skills[(test1>0), ]
test2 <- rep(0, nrow(skills))
for(i in 1:nrow(skills)) test2[i] <- sum(is.na(skills[i,]))
skills <- skills[test2==0,]
scaledSkills <- skills
for(i in 1:nrow(skills)) scaledSkills[i, ] <- rank(skills[i, ])
op <- par(mar = c(12,4,4,2) + 0.1)
boxplot(scaledSkills, las=2)

cm <- cor(scaledSkills, use="complete.obs", method="pearson")

library(ggplot2)
library(reshape2)
ggplot(melt(cm), aes(Var1, Var2, fill=value)) + geom_tile() +
		scale_fill_gradient2(limits=c(-1, 1)) +
		theme(axis.text.x=element_text(angle=-90, hjust=0))

# NMF stuff moved to separate file

d <- dist(scaledSkills)
h <- hclust(d)
plot(h)
dsScore <- apply(scaledSkills, 1, sum) # all sum to 253 (=sum(1:22)), as wanted

p <- prcomp(scaledSkills)
biplot(p, xlabs=rep(".", nrow(scaledSkills)))
screeplot(p)
scree <- p$sdev^2; names(scree) <- 1:length(scree)
barplot(scree, xlab="Principal component", ylab="variance explained")
cumscree <- cumsum(scree); barplot(cumscree/cumscree[length(cumscree)])

d <- dist(t(scaledSkills))
h <- hclust(d)
op <- par(mar = c(8,4,4,2) + 0.1)
plot(h, xlab="")

skillScore <- apply(scaledSkills, 2, sum)/(nrow(scaledSkills)*ncol(scaledSkills))
op <- par(mar = c(12,4,4,2) + 0.1)
barplot(skillScore[order(skillScore)], las=2)
#axis(1, at=1:22, labels=names(scaledSkills), las=2)
par(op)
```

## Data Scientist types
```{r types}
# names(mydata)[55:65]
types <- mydata[, names(mydata)[grepl("^Type", names(mydata)) & !grepl("Other$", names(mydata))]]
for(i in 1:ncol(types)) {
	types[, i] <- as.character(types[, i])
	types[types[, i]=="", i] <- "0" 
	types[types[, i]=="Maybe", i] <- "3" 
	types[types[, i]=="Maybe a bit", i] <- "2" 
	types[types[, i]=="Not at all", i] <- "1" 
	types[types[, i]=="Yes, but not completely", i] <- "4" 
	types[types[, i]=="Yes, definitely", i] <- "5"
	types[, i] <- as.numeric(types[, i])
}
colnames(types) <- sub("^Type", "", names(types))
colnames(types)[1] <- "Developer"
colnames(types)[5] <- "Statistician"
colnames(types)[6] <- "Jack-of-all-trades"
test1 <- apply(types, 1, var)
types <- types[(test1>0), ]
test2 <- apply(types, 1, sum)
types <- types[(test2>0), ]
scaledTypes <- types
for(i in 1:nrow(types)) scaledTypes[i, ] <- rank(types[i, ])
op <- par(mar = c(8,4,4,2) + 0.1)
boxplot(scaledTypes, las=2)

cm <- cor(scaledTypes, use="complete.obs", method="pearson")

ggplot(melt(cm), aes(Var1, Var2, fill=value)) + geom_tile() +
		scale_fill_gradient2(limits=c(-1, 1)) +
		theme(axis.text.x=element_text(angle=-90, hjust=0))

# NMF moved to separate .Rmd file

typeScore <- apply(scaledTypes, 2, sum)/(nrow(scaledTypes)*ncol(scaledTypes))
op <- par(mar = c(8,4,4,2) + 0.1)
barplot(typeScore[order(typeScore)], las=2)
#axis(1, at=1:22, labels=names(scaledTypes), las=2)
par(op)

d <- dist(scaledTypes)
h <- hclust(d)
op <- par(mar = c(8,4,4,2) + 0.1)
plot(h, xlab="")

d <- dist(t(scaledTypes))
h <- hclust(d)
op <- par(mar = c(8,4,4,2) + 0.1)
plot(h, xlab="")
```

## Motivation to attend meetup evenings
```{r motivation}
#names(mydata[, 67:75])
motivationNames <- names(mydata[, 67:75])
motivationCounts <- rep(0, length(motivationNames))
for(i in 1:length(motivationNames)) motivationCounts[i] <- sum(mydata[, motivationNames[i]]!="")
motivationDisplayNames <- c("Doesn't attend", "Socialising", "Business to Business", "Meeting employers", "Meeting employees", "Developments global", "Developments Belgium", "Contributing", "Other")
names(motivationCounts) <- motivationDisplayNames
op <- par(mar = c(12,4,4,2) + 0.1)
barplot(motivationCounts, las=2)
responded <- sum(completed[, which(colnames(completed)=="Motivation")]); skipped <- nrow(mydata)-responded
```

## Activities
```{r activities}
#names(mydata)[77:82]
scaledActivities <- mydata[, 77:81]
for(i in 1:ncol(scaledActivities)) {
	scaledActivities[, i] <- as.character(scaledActivities[, i])
	scaledActivities[scaledActivities[, i]=="", i] <- "0" 
	scaledActivities[scaledActivities[, i]=="No, not our role", i] <- "1" 
	scaledActivities[scaledActivities[, i]=="No opinion either way", i] <- "3" 
	scaledActivities[scaledActivities[, i]=="Probably not our role", i] <- "2" 
	scaledActivities[scaledActivities[, i]=="Yes, probably", i] <- "4" 
	scaledActivities[scaledActivities[, i]=="Yes, definitely", i] <- "5"
	scaledActivities[, i] <- as.numeric(scaledActivities[, i])
}
colnames(scaledActivities) <- sub("^Activities", "", names(scaledActivities))
test1 <- apply(scaledActivities, 1, var)
scaledActivities <- scaledActivities[(test1>0), ]
test2 <- apply(scaledActivities, 1, sum)
scaledActivities <- scaledActivities[(test2>0), ]
for(i in 1:nrow(scaledActivities)) scaledActivities[i, ] <- rank(scaledActivities[i, ])
op <- par(mar = c(8,4,4,2) + 0.1)
boxplot(scaledActivities, las=2)
responded <- nrow(scaledActivities)
skipped <- nrow(mydata)-responses

activityScore <- apply(scaledActivities, 2, sum)/(nrow(scaledActivities)*ncol(scaledActivities))
op <- par(mar = c(8,4,4,2) + 0.1)
barplot(activityScore[order(activityScore)], las=2)
```

## Training
```{r training}
# names(mydata)[83:87]
scaledTraining <- mydata[, 83:87]
for(i in 1:ncol(scaledTraining)) {
	scaledTraining[, i] <- as.character(scaledTraining[, i])
	scaledTraining[scaledTraining[, i]=="", i] <- "0" 
	scaledTraining[scaledTraining[, i]=="No, not our role", i] <- "1" 
	scaledTraining[scaledTraining[, i]=="No opinion either way", i] <- "3" 
	scaledTraining[scaledTraining[, i]=="Probably not our role", i] <- "2" 
	scaledTraining[scaledTraining[, i]=="Yes, probably", i] <- "4" 
	scaledTraining[scaledTraining[, i]=="Yes, definitely", i] <- "5"
	scaledTraining[, i] <- as.numeric(scaledTraining[, i])
}
colnames(scaledTraining) <- sub("^Training", "", names(scaledTraining))
test1 <- apply(scaledTraining, 1, var)
scaledTraining <- scaledTraining[(test1>0), ]
test2 <- apply(scaledTraining, 1, sum)
scaledTraining <- scaledTraining[(test2>0), ]
for(i in 1:nrow(scaledTraining)) scaledTraining[i, ] <- rank(scaledTraining[i, ])
op <- par(mar = c(8,4,4,2) + 0.1)
boxplot(scaledTraining, las=2)
responded <- nrow(scaledTraining)
skipped <- nrow(mydata)-responded

trainingScore <- apply(scaledTraining, 2, sum)/(nrow(scaledTraining)*ncol(scaledTraining))
op <- par(mar = c(8,4,4,2) + 0.1)
barplot(trainingScore[order(trainingScore)], las=2)

```